Computer-Assisted Analysis of the \\ Anderson-Hájek Ontological Controversy

### 2015-07-13 Papers/InProgress/HajekAndersonControversy/main.tex
Resolve inconsistencies between result table and text.

Although Hájek did not notice the superfluousness of A4 and A5' in his
\AOEH, he did describe yet another emendation (his \AOEHH, cf.
Fig. \ref{fig:Hajek2}) where A4 and A5' are superfluous (though no
claim is made w.r.t. to their redundancy), if A3' is replaced by a
stronger axiom (H:A3) additionally stating that the property of actual
existence is positive when it comes to God-like beings
\cite[sec.~5]{Hajek2002}. Formalization of \AOEHH\ shows that A4 is
not only superflous, but also redundant. For A5', no conclusive results
were achieved; neither a proof nor a countermodel could be
automatically generated. Surprisingly, a countermodel for the weaker
A3' was succesfully generated. This is somewhat unsatisfactory (for
theistic goals), because it shows that \AOEHH\ does not entail the
positiveness of being God-like.

@@

The formalization of \AOEHHH\ shows that H:A4 is
independent. The old A5' is independent as well, and both H:A4 and H:A5
are superfluous, but no conclusive results were achieved regarding
independence or redundancy of H:A5.

Additionally, Anderson \citep[footnote
14]{anderson90:_some_emend_of_goedel_ontol_proof} remarks that only
the quantifiers in T3' and in A:D2 need to be interpreted as
actualistic quantifiers, while others may be taken as possibilistic
quantifiers. Our computer-assisted study of this mixed variant shows
that A4 is still redundant, but A5' becomes independent (hence not
redundant). Unfortunately, a countermodel for T3 can then be found.

@@

He takes the property of being God-like as a primitive and defines (B:D1) the positive properties as
those properties necessarily possessed by every God-like being. He
then briefly indicates (B:L1) that B:D1 is logically equivalent, under
modal logic \SFour, to the conjunction of D1', A2', A3' and A4'. This
has been confirmed in the computer-assisted formalization: A2' and A3'
can be quickly automatically derived in logic \K. A4' can be proved in
logic \KT (i.e. assuming reflexivity of the accessibility relation).
For constant domain semantics, proving D1' is possible in logic \KFour,
whereas for varying domain semantics, a countermodel can be found
even in logic \SFive. Conversely, the proof that B:D1 is entailed by
D1', A2', A3' and A4' is possible already in logic \K. The provers
also show that theorem T3' follows from B:D1, B:D2, B:D3, A:A1' and
A5' already in logic \KB. Bjordal's last paragraph briefly mentions
Hájek's ideas about the superfluousness of A5' and claims that it is
possible, with (unclear) additional modifications of the definitions,
to eliminate A5' from his theory as well. Without any additional
modification, the automated reasoners show that A5' is actually not
superfluous. All these results, with the exception of the
aforementioned countermodel for D1', hold for both constant and
varying domain semantics.

>>>>

Although Hájek did not notice the superfluousness of A4 and A5' in his
\AOEH, he did describe yet another emendation (his \AOEHH, cf.
Fig. \ref{fig:Hajek2}) where A4 and A5' are superfluous (though no
claim is made w.r.t. to their redundancy), if A3' is replaced by a
stronger axiom (H:A3) additionally stating that the property of actual
existence is positive when it comes to God-like beings
\cite[sec.~5]{Hajek2002}. Formalization of \AOEHH\ shows that A4 is
not only superflous, but also redundant. In addition, A5' becomes superfluous
and independent. Surprisingly, a countermodel for the weaker
A3' was succesfully generated. This is somewhat unsatisfactory (for
theistic goals), because it shows that \AOEHH\ does not entail the
positiveness of being God-like.

@@

The formalization of \AOEHHH\ shows that both H:A4 and H:A5 are superfluous
aswell as independent. For the old A5', no conclusive results were achieved.

Additionally, Anderson \citep[footnote
14]{anderson90:_some_emend_of_goedel_ontol_proof} remarks that only
the quantifiers in T3' and in A:D2 need to be interpreted as
actualistic quantifiers, while others may be taken as possibilistic
quantifiers. Our computer-assisted study of this mixed variant shows
that A4 is still redundant in logic \KFourB, but A5' becomes independent (hence not
redundant). Unfortunately, a countermodel for T3 can then be found.

@@

He takes the property of being God-like as a primitive and defines (B:D1) the positive properties as
those properties necessarily possessed by every God-like being. He
then briefly indicates (B:L1) that B:D1 is logically equivalent, under
modal logic \SFour, to the conjunction of D1', A2', A3' and A4'. This
has been confirmed in the computer-assisted formalization: A2' and A3'
can be quickly automatically derived in logic \K. A4' can be proved in
logic \KT (i.e. assuming reflexivity of the accessibility relation).
For constant domain semantics, proving D1' is possible in logic \KFour,
whereas for varying domain semantics, a countermodel can be found
even in logic \SFive. Conversely, the proof that B:D1 is entailed by
D1', A2', A3' and A4' is possible already in logic \K. The provers
also show that theorem T3' follows from B:D1, B:D2, B:D3, A:A1' and
A5' already in logic \KB. Bjordal's last paragraph briefly mentions
Hájek's ideas about the superfluousness of A5' and claims that it is
possible, with (unclear) additional modifications of the definitions,
to eliminate A5' from his theory as well. Without any additional
modification, the automated reasoners show that A5' is independent, but actually
not superfluous. All these results, with the exception of the
aforementioned countermodel for D1', hold for both constant and
varying domain semantics.
